---
title: "Exercises"
author: "Simon Parker"
date: "19 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

```{r}
a = 1
```

Did anything happen? We can check by typing in `a` to see if `R` returns the value `1`.

```{r}
a # Click "Run" or the green triangle or hit "Ctrl + Enter"
```

Now, since `a` is a *variable*, we can replace its value by assigning it a new value.

```{r}
a = "Boo!"
a # 'a' no longer returns the value 1.
```

#Exercise 2
### Vectors
To create a vector, we can `c`ombine or `c`oncatenate values as follows:
```{r}
## Numeric Vector
c(1, 2)
1:2
```
```{r}
## Character Vector
c("Alice", "Bob", "Cat")
c("A", "B", "C")
LETTERS[c(1, 2, 3)]
LETTERS[1:3]
letters[1:3]
```
```{r}
## Logical Vector
c(TRUE, FALSE)
c(1, 2)[c(TRUE, FALSE)]
```


## Question: Explain what is going on in the code below:

```{r}
c(1, "A") 
c(1, 2, 3, 4)[c(TRUE, FALSE)]
c(1, 2, 3, 4)[c(TRUE, FALSE, TRUE, FALSE)]

```

### Matrices
```{r}
## Numeric Matrix
i = 1:6
matrix(i, nrow = 3)
matrix(i, nrow = 2, ncol = 3)
matrix(i, ncol = 2)
cbind(1:3, 4:6)                  # column bind
rbind(c(1, 4), c(2, 5), c(3, 6)) # row bind
```
```{r}
## Character Matrix
matrix(LETTERS[i], nrow = 3)
matrix(i, nc = 2, byrow = TRUE)
```

### Lists
```{r}
x1 = list(1, "Alice")
x1
x2 = list(TRUE, 2, "Bob")
x2
x3 = list(x1, x2)
x3
```

### Data Frames
```{r}
x1 = 1:3
x2 = LETTERS[x1]
x3 = data.frame(id = x1, cat = x2)
x3
data.frame(x3, x3)
data.frame(x3, x3, x3)
data.frame(list(a = x1, b = x2))
```

### Checking Types
```{r}
is.numeric(1:10)
is.numeric(c("A", "B"))
is.data.frame(matrix(1:10, nrow = 2))
is.matrix(data.frame(1:10, 2:11))
is.list(matrix(1:10, ncol = 2))
```

## Question: is a data.frame a list?



## Question: is a list a data.frame?



# Exercise 3

```{r}

greet_student = function(name) {
  print(paste0("Hello, ",
              name,
              "! ",
              "Welcome to Encounters with Big Data 2018!"))
}

```
```{r}

student_names = c("Simon", "Deborah", "Myles", "Louise", "Jason Isaacs")
```
```{r}
sapply(student_names, greet_student)
```

# Exercise 4

```{r}
f = function(n) {
  `*` = `/`
  n * n
}

f(2)

2 * 2
```

# Exercise 5

## Memory limits

```{r}
memory.limit()
```
Let's try and hit our memory limit by simulating a large matrix.
```{r}
## Simulate a large matrix.
zero_mat = matrix(0, 1e7, ncol = 10)
dim(zero_mat)
print(object.size(zero_mat), unit = "Gb")
print(str(zero_mat))
rm(zero_mat)
```

This time, we simulate a large matrix of *integer* values.
```{r}
## Simulate a large integer matrix.
int_mat = matrix(as.integer(0), 1e+7, ncol = 10)
dim(int_mat)
print(object.size(int_mat), unit = "Gb")
print(str(int_mat))
rm(int_mat)
```

### Why do these matrices have the same dimensions but have different file sizes? 

## Copying
```{r}
## We create a data frame consisting of 2 million random integers.
## Imagine that these are counts the number of steps people have
## taken across two days.
day1 = sample(9000:20000, 1e6, replace = TRUE)
day2 = sample(9000:20000, 1e6, replace = TRUE)
steps = data.frame(day1, day2)
head(steps)

## Suppose that there was a fault in the collection device and we need
## to increment the first 1000 counts for both days.
m = 1000

## Data frame way.
x <- system.time({
  for (i in 1:m) {
    steps[i, ] = steps[i, ] + 1
  }
})

## Vector way. This offloads the update process down to C code, so
## the copying is under control.
y <- system.time({
  for (i in 1:m) {
    day1[i] = day1[i] + 1
    day2[i] = day2[i] + 1
  }
  steps2 = data.frame(day1, day2)
})

identical(steps, steps2)

x
y

```

# Exercise 6

Create a dataset using the nycflights13 data that only contains columns that start with 'arr' and end with 'delay'.

Hint: You may need to 'library' dplyr and nycflights13 first; the "starts_with" and "ends_with" commands may also be helpful

```{r}

```

# Exercise 7

```{r}

```

# Exercise 8

```{r}

```

# Exercise 10

```{r}

```

Spark demo

```{r}
library(sparklyr)
library(dplyr)
library(nycflights13)
library(ggplot2)

sc <- spark_connect(master="local")
flights <- copy_to(sc, flights, "flights", overwrite = T)
airlines <- copy_to(sc, airlines, "airlines", overwrite = T)
planes <- copy_to(sc, planes, "planes", overwrite = T)
weather <- copy_to(sc, weather, "weather", overwrite= T)

src_tbls(sc)
```

Now our flights and airlines data can be worked on, without being copied to memory. This makes it possible to work with large datasets.

We can then use dplyr commands that we have already seen without using up our memory.
```{r}
flights %>%
  inner_join(airlines, by = "carrier") %>%
  inner_join(planes, by = "tailnum")
  
```

When we want to work on the data for plotting and analysis we have to use the collect() command. This will bring the data into R.

```{r}
cleandata = flights %>% 
  select(dest, dep_delay, month, day) %>%
  na.omit() %>%
  collect()
```

Now we have a cleaned datset we can start working on.